---
title: "Data-tidying-lesson"
author: "Anna Talucci"
date: "October 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
```
To access `filter` from the `stats` package:
`stats::filter()`

# Read in and clean up data

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"), stringsAsFactors = FALSE)
head(catch_original)
```



Select the columns we want using `select()`

```{r}
catch_data = catch_original %>%
  # select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
  # select(-All, -notesRegCode)
  select(Region:Chum)

head(catch_data)
```

```{r}
summary(catch_data)
```

Change values in the Chinook column to numeric using `mutate()`

```{r}
catch_clean = catch_data %>% 
  mutate(Chinook=as.numeric(Chinook))

head(catch_clean)
```

investigate using `which()` and `is.na()`

```{r}
i = which(is.na(catch_clean$Chinook))
i
```

Use [] to look at individual rows and columns [row, column]

```{r}
catch_original[i,]
```

* Use `mutate()` to change the I to a 1
* Use `mutate()` to coerce Chinook Column to numeric

`|` serves as the or operator ifelse(Chinook=="I" | Chinook == "l", 1, Chinook
`%in%` also serves as or operator

```{r}
catch_clean = catch_data %>% 
  mutate(Chinook = ifelse(Chinook %in% c("I", "l"), 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```


# Tidy data

Move from wide format to tall format using `pivot_longer()`

```{r}
catch_long = catch_clean %>% 
  pivot_longer(cols = -c(Region, Year),
               names_to = "species",
               values_to ="catch")
head(catch_long)
```


```{r}
catch_wide = catch_long %>% 
  pivot_wider(names_from = species,
              values_from = catch)

head(catch_wide)
```

`rename()` catch to catch_thousands
`mutate()` to create a new catch column

```{r}
# catch_long = catch_long %>% 
#   rename(catch_thousands = catch) %>% 
#   mutate(catch = catch_thousands * 1000) %>% 
#   select(-catch_thousands)

catch_long = catch_long %>% 
  mutate(catch = catch * 1000)
  
head(catch_long)  
```


# Summarize data

mean catch by region and species 

```{r}
mean_region = catch_long %>% 
  group_by(Region, species) %>% 
  summarise(mean_catch = mean(catch), n_obs = n())

mean_region
```

mean sockeye  catch by region
```{r}
mean_sockeye_region = catch_long %>% 
  filter(species == "Sockeye") %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))

mean_sockeye_region
```

```{r}
annual_catch = catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(desc(total_catch))

summary(annual_catch)
```


```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = FALSE) %>% 
  select(code, mgmtArea)



head(region_defs)
```

```{r}
catch_joined = left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)
```


```{r}
region_defs = region_defs %>% 
  rename(Region = code)

catch_joined = left_join(catch_long, region_defs, by = "Region")

head(catch_joined)
```

# Separate and Unite Example

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE) %>%
  unite(date2, year, month, day, sep = "-")

dates_df %>% 
  mutate(three_char = stringr::str_sub(date, start = -4, end = -1))
```

lumridate package for managing and manipulating dates 
**THE END**
